---
title: "BayLum: An Introduction"
author: "Claire Christophe, Anne Philippe, Sebastian Kreutzer, Guillaume Guerin"
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette:
    self-contained: yes
    standalone: yes
    toc: yes
vignette: >
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment = "")
options(width = 120, max.print = 100)
library(BayLum)
```

# Introduction

BayLum provides a collection of various R functions for Bayesian analysis of Luminescence data. This includes, amongst others, data import, export, application of age models and palaeodose model.

It is possible to process data for various samples simultaneously and to consider more than one BIN file per sample.
Single-grain and Multi-grain OSL measurement can be analysed.
Stratigraphic constrains and systematic errors can be added in the analyse.

For those who already know how to use R, BayLum won't be difficult to use. 

### Installing *BayLum* package

To install BayLum you first need to install [R](https://www.r-project.org) and  [Rstudio](http://www.rstudio.com) that has a nice desktop environment for using R. In addition, you need to install [JAGS](http://mcmc-jags.sourceforge.net) for the Bayesian analysis of the models. Once in R (or in RStudio) you can type:

```{r,eval=FALSE}
install.packages('BayLum')
```

at the R command prompt to install BayLum If you then type:

```{r}
library(BayLum)
```

it will load in all the BayLum functions.

# 1. Age analysis of one sample

## 1.1. Generate data file from BIN file of Single-grain OSL measurements

A first step is to read information from BIN file.

Let us consider the sample named *GDB3*. 
All informations concerning this sample (environmental dose, source dose, disc and position of grains,... see "What are the required files in each subfolder?" section in help of  **Generate\_DataFile** function) are containing in a subfolder named as the sample *GDB3* located in *path*.

Then, you can use the function **Generate\_DataFile** as follows:

```{r,eval=FALSE}
path="path" # Put path in quotes to go to the folder containing bin.BIN file and associated .csv files
            # Note that in parameter "Path", the character must be terminated by "/"
folder=c("GDB3")# give the name of the data folder
nbsample=1       # give the number of samples
DATA1=Generate_DataFile(Path=path,FolderNames=folder,Nb_sample=nbsample)
```

##### Remark 1: save en load information
The user can use the function **save** to save output $DATA1$ as follows:
```{r,eval=FALSE}
save(DATA1,file=c(paste(path,"DATA1.RData",sep="")))
```
The user can acced to this file with the function **load**:
```{r,eval=FALSE}
load(file=c(paste(path,"DATA1.RData",sep="")))
```

##### Remark 2: general structure of the data
To check if there is no problem, the user can read information on BIN file and associated .csv files as folows:

```{r, echo=FALSE}
data("DATA1") # load output of Generate_DataFile function for the sample named "GDB3"
```

```{r}
str(DATA1)
```

In *DATA1*, there is information for the sample named *GDB3*.
It is expected (for more information we refer to "Value" section in help of  **Generate\_DataFile** function):

  - 1 list (1 list per sample), for DATA\$LT, DATA\$sLT, DATA1\$ITimes and DATA1\$regDose}
  - 1 vector of size 2 (a 2 column matrix, with 1 line per sample), for DATA1\$dLab, DATA1\$ddot_env
  - 1 integer (1 integer per BIN files), for DATA1\$J, DATA1\$K, DATA1\$Nb_measurement.

##### Remark 3: Lx/Ts value as a function of Regenerative Dose
You can also see the Lx/Tx values as a function of regenerative dose using the function **LT\_RegenDose**:

```{r,eval=FALSE,fig.cap = "inst/LT_RegenDose_plot.pdf"}
LT_RegenDose(DATA=DATA1,Path=path,FolderNames=folder,
             SampleNames=folder,Nb_sample=nbsample)
```
Note that, here we consider only one sample, and the name of the folder is the name of the sample, that is why $FolderNames=$ *folder* and $SampleNames=$ *folder*. 

## 1.2. Generate data file from BIN file of Multi-grain OSL measurements

For a multi-grain OSL measurements, instead of **Generate\_DataFile** function, the user use **Generate\_DataFile_MG** whith similar parameter. 
The difference come from the associated *Disc.csv* file instead *DiscPos.csv* file for Single-grain OSL Measurements (we refer to help of **Generate\_DataFile_MG** function).

The user can also plot Lx/Tx values as a function of Regenerative Dose, for every selected aliquot and for each sample, with the **LT_RegenDose** with the option $SG=$*False*.

## 1.3. Age analysis

To compute age of the sample *GDB3*, you can run the following code:

```{r,eval=FALSE}
priorage=c(10,100)    # GDB3 is an old sample
Age=Age_Computation(DATA=DATA1,SampleName="GDB3",
                    PriorAge=priorage,
                    distribution="cauchy", # Option on equivalent dose distribution
                    LIN_fit=TRUE,Origin_fit=FALSE, # Option on growth curves
                    Taille=10000)
```

If $DATA1$ is output of **Generate\_DataFile_MG**, then the user can use **Age_Computation** exactly as in the previous example.

##### Remark 0: on MCMC trajectories

If MCMC trajectories did not converge, you can add more iteration at the parameter $Taille$ in the function **Age_Computation**, for example $Taille=20000$ or $Taille=50000$.

Or you can precise the prior distribution if it was not spécified before (for example if it is a young sample $PriorAge=$ *c(0.01,10)*, if it is an old sample $PriorAge=$*c(10,100)*).

If trajectories are still not converged, you can check if the choice of $distribution$ and growth cruves are good.

##### Remark 1: $LIN\_fit$ and $Origin\_fit$, dose response curves option

By default, a saturating exponential plus linear growth curve is considered, but the user can choose other growth curves by changing option $LIN\_fit$ and $Origin\_fit$ in the function. We refer to the user manual for more detail. 

What we can recommend is to use ... if you have...


##### Remark 2: $distribution$, equivalent dose dispersion option

By default, a *cauchy* distribution is considered, but the user can choose other distribution replacing *cauchy* by *gaussian* or *lognormal_A* or *lognormal_M* in $distribution$ option.  

The difference between the models: *lognormal_A* and *lognormal_M* is that the equivalent dose dispersion are distributed according to:

 - a lognormal distribution with mean or Average equal to the palaeodose for the first model
 - a lognormal distribution with Median equal to the palaeodose for the second model. 

If you are in this cases: ..... , you can consider .....

##### Remark 3: $SavePdf$ and $SaveEstimates$ option

These two options allow to save directly information. 

$SavePdf=TRUE$ saves a pdf file with MCMC trajectories of parameters $A$ (age), $D$ (palaeodose), $sD$ (equivalent doses dispersion). The user must precise $OutputFileName$ and $OutputFilePath$ to define name and path of the pdf file. 

$SaveEstimates=TRUE$ saves a csv file containing Bayes estimates, credible interval at 68\% and 95\%  and the Gelman and Rudin test of convergency of the parameters $A$, $D$, $sD$. 
The user must precise $OutputTableName$ and $OutputTablePath$ to define name and path of the csv file.

##### Remark 4: $PriorAge$ option

By default, an age between 0.01 Ka and 100 ka is expected. 
If the user has more informations on the sample age it can change $PriorAge$.

For example, if you know that the sample is an old sample, you can choose $PriorAge=c(10,120)$. In contrast, if you know that is a young sample, you can choose $PriorAge=c(0.001,10)$. Note that, it is not possible to consider as lower bound 0, you can consider 0.001. 

You must take care with this option, because $PriorAge$ are the lower bound and the upper bound of the estimated age. If you give a $PriorAge$ too precise it can bias the result. 


## 1.4. Various BIN files for one sample

In the previous example we consider the samplest case: one sample, and one BIN file for this sample.
But we can consider various BIN files for one sample.
To do this, you write in $Names$ the names of subfolder coresponding to a specific BIN file, both located in $path$. 
For example $Names=c("name\_binfile1","name\_binfile2")$.
Then, you can complete **Generate_DataFile** (or **Generate_DataFile_MF**) as follows:

```{r,eval=FALSE}
nbsample=1
nbbinfile = length(Names)
Binpersample = c(length(Names))
DATA_BF=Generate_DataFile(Path=path,
                          FolderNames=Names,
                          Nb_sample=nbsample,
                          Nb_binfile=nbbinfile,
                          BinPerSample=Binpersample)
# Computation of age
Age=Age_Computation(DATA=DATA_BF,
                    SampleName="Names",
                    BinPerSample=Binpersample)
```

# 2. Age analysis of various samples

## 2.1. Generate data file from BIN files

More precisely, the function **Generate\_DataFile**  (or **Generate_DataFile_MF**) can consider various samples simultaneously, and can consider more than one BIN file per sample.

Assume that we are interested in two samples named: *sample1* and *sample2*. In addition, we have two BIN files for the first sample named: *sample1-1* and *sample1-2*, and one Bin file for the sample2, named *sample2-1*.
Then we must have 3 subfolders named *sample1-1*, *sample1-2* and *sample2-1*; which each subfolder contains one BIN file named **bin.BIN**, and its associated files **DiscPos.csv**, **DoseEnv.csv**, **DoseSourve.csv** and **rule.csv**. These 3 subfolders are located in *path*.

To fill corectly $BinPerSample$: 
$$binpersample=c(\underbrace{2}_{\text{sample 1: 2 bin files}},\underbrace{1}_{\text{sample 2: 1 bin file}})$$

```{r,eval=FALSE}
Names=c("sample1-1","sample1-2","sample2-1") # give the name of the folder datat
nbsample=2    # give the number of samples
nbbinfile=3   # give the number of bin files
DATA=Generate_DataFile(Path=path,FolderNames=Names,
                       Nb_sample=nbsample,
                       Nb_binfile=nbbinfile,
                       BinPerSample=binpersample)
```

#### **Concat_DataFile** function

If the user has already saved informations obtained with **Generate_DataFile** function  (or **Generate_DataFile_MG** function) about samples in RData file. The user can concatenate data with the **Concat_DataFile**. 

For example if *DATA1* is output of sample named "GDB3", and *DATA2* is output of "GDB5",

```{r}
data("DATA1",envir = environment())
data("DATA2",envir = environment())
DATA3=Concat_DataFile(L1=DATA2,L2=DATA1)
str(DATA3)
```

In DATA, there is information for samples named *GDB3* and *GDB5*, then it is expected 

  - 2 lists (1 list per sample) for  DATA\$LT, DATA\$sLT, DATA1\$ITimes and DATA1\$regDose
  - a matrix with 2 columns (1 line per sample) for DATA1\$dLab, DATA1\$ddot_env
  - 2 integers (1 integer per BIN files here we have 1 BIN file per sample) for DATA1\$J, DATA1\$K, DATA1\$Nb_measurement.

It is possible to concat data from Single-grain OSL measurement and from Multi-grain OSL measurement.

You can also see the L/T in function of the regenerative doses for each sample as follows:

```{r,eval=FALSE}
LT_RegenDose(DATA=DATA3,Path=path,FolderNames=Names,
             Nb_sample=nbsample,SG=rep(TRUE,nbsample))
```

As in *DATA3* there are informations concerning sample *GDB3* and *GDB5* which are Single-grain OSL measurements, $SG[1]=SG[2]=TRUE$. If one sample (whose number ID is equal to $i$) is Multi-grain OSL measurements, the user changes $SG$ option by $SG[i]=$*FALSE*.

## 2.2. Age analysis without constraint

If there is no stratigraphic constraint, you can compute the following code to analyse simultaneously the age of the sample *GDB5* and *GDB3*.

```{r,eval=FALSE}
priorage=c(1,10,10,100) # see remark to have more information on this parameter
Age=AgeS_Computation(DATA=DATA3,Nb_sample=2,SampleNames=c("GDB5","GDB3"),
              PriorAge=priorage,
              distribution="cauchy", # Option on equivalent dose distribution
              LIN_fit=TRUE,Origin_fit=FALSE, # Option on growth curves
              Iter=10000)
```

##### Remark: MCMC trajectories and convergencies
If MCMC trajectories did not converge, we refer to remark 0 in Section 1.3. 

##### Remark
Options detailled in remark 1, 2, 3, of the section1.3, are still available for this function.

As for the function **Age_computation**, age for each sample is by default between 0.01 Ka and 100 ka. 
If you have more informations on your samples it is possible to change $PriorAge$ parameters. 
$PriorAge$ is a vector of size = 2*$Nb\_sample$, the two first values of $PriorAge$ concern the first sample and so on. 

For example, if you know that sample named *GDB5* is a young sample whose its age is between 0.01 ka and 10 ka, and *GDB3* is an old sample whose age is between 10 ka and 100 ka, 
$$PriorAge=c(\underbrace{0.01,10}_{GDB5\ prior\ age},\underbrace{10,100}_{GDB3\ prior\ age})$$


## 2.3. Stratigraphic constraints

With the function **AgeS_Computation** it is possible to take into account stratigraphic relations between samples.

For example, we know that *GDB5* age is supposed lower than *GDB3* age.

#### a) Ordering samples

To take into acount stratigraphic constraints, informations on samples must be ordered. 

Either you inter sample Name (corresponding to subforlder names) in $Names$ parameter of the function **Generate_DataFile**, ordered by order of increasing ages; 
or you enter saved .RData informations of each sample in **Concat_DataFile**, ordered by order of increasing ages.

```{r,eval=FALSE}
# using Generate_DataFile function
Names=c("GDB5","GDB3")
nbsample=2
Generate_DataFile(Path=path,FolderNames=Names,Nb_sample=nbsample)

# using Concat_DataFile function
data(DATA1,envir = environment()) # .RData on sample GDB3
data(DATA2,envir = environment()) # .RData on sample GDB5
DATA3=Concat_DataFile(L1=DATA1,L2=DATA2)
```


#### b) Construct stratigraphic constraint matrix

Let *SC* be the matrix containing all informations on stratigraphic relations for this two samples.
This matrix is defined as follows:

  - the size of the matrix: row number of $StratiConstraints$ matrix is equal to $Nb\_sample+1$, and column number is equal to $Nb\_sample$.

  - the first line of the matrix: for all i in {1,...,$Nb\_Sample$}, $StratiConstraints[1,i]=1$, that means the lower bound of the sample age given in $PriorAge[2i-1]$ for the sample whose number ID is equal to $i$, is taken into account.

  - the sample relations: for all j in {2,...,$Nb\_Sample$+1} and all i in {j,...,$Nb\_Sample$}, $StratiConstraints[j,i]=1$ if sample age whose number ID is equal to j-1 is lower than sample age whose number ID is equal to i. Otherwise, $StratiConstraints[j,i]=0$.

The user can use the function *SCMatrix* to define this matrix:

```{r}
SC=SCMatrix(Nb_sample=2,SampleNames=c("GDB5","GDB3"))
1
```

In our case: 2 samples, *SC* is a matrix with 3 lines and 2 columns. The first line containts $c(1,1)$ (because we take into acount prior ages), the second line containts $c(0,1)$ (because the sample 2, named *GDB3* is supposed older than the sample 1, named *GDB5*) and the third line contains $c(0,0)$ (because the sample 2, named *GDB3* is not younger than the sample 1, named *GDB5*). 
Then, we can also fill the matrix of stratigraphic relations as follow:

```{r,eval=TRUE}
(SC=matrix(data=c(1,1,0,1,0,0),ncol=2,nrow = (2+1),byrow = T))
```

#### c) Age computation

```{r,eval=FALSE}
Age=AgeS_Computation(DATA=DATA3,Nb_sample=2,SampleNames=c("GDB5","GDB3"),
                   PriorAge=priorage,
                   distribution="cauchy", # Option on equivalent dose distribution
                   LIN_fit=TRUE,Origin_fit=FALSE, # Option on growth curves
                   StratiConstraints=SC,  # Stratigraphic relations 
                   Taille=10000)
```



# References

Models used in this package are detailled in the following publication:

  - Combes, B., Philippe, A., Lanos, P., Mercier, N., Tribolo, C., Guerin, G., Guibert, P., Lahaye, C., (2015). A Bayesian central equivalent dose model for optically stimulated luminescence dating. Quaternary Geochronology 28, 62-70.
  - Combes, B. and Philippe, A. (2017). Bayesian analysis of multiplicative Gaussian error for multiple ages estimation in optically stimulated luminescence dating. Quaternary Geochronology (in press)

For more details on the diagnostic of Markov chain :

  - Robert and Casella (2009). Introducing Monte Carlo Methods with R. Springer Science & Business Media.

For more detail on data used in examples, we refer to the following publication:

  - Tribolo, C., Asrat, A., Bahain, J. J., Chapon, C., Douville, E., Fragnol, C., Hernandez, M., Hovers, E., Leplongeon, A.,  Martin, L,  Pleurdeau, D,  Pearson, O ,  Puaud, S,  Assefa, Z. (2017). Across the Gap: Geochronological and Sedimentological Analyses from the Late Pleistocene-Holocene Sequence of Goda Buticha, Southeastern Ethiopia. PloS one, 12(1), e0169418.
